Tree ring bayesian hierarchical mixed effects model
========================================================

The tree ring analysis to-date has included a hierarchical structure to include individuals, subplots, plots and sites and some individual effects like species. I am now adding year x plot and year x site interactions to see how much variance is explained by climate at each site and within a site.

You can embed an R code chunk like this:

```{r}
#Required Packages
library(lme4)
library(rjags)
library(data.table)
library(mcmcplots)

#Source required PEcAn scripts for reading and matching treerings and plot data
source("/Users/Josh/Documents/Research/pecan/modules/data.land/R/Read_Tuscon.R")
source("/Users/Josh/Documents/Research/pecan/modules/data.land/R/matchInventoryRings.R")
```

Using most up-to-date version of adult census data for treering matching. Includes 2015 census data, but not boolean damage column. All other columns should be the same as analysis done for AGU and EE seminar.

```{r}
#read Tucson format ringwidth files
trees.diss <- read.csv("/Users/josh/Dropbox/Dissertation/CH1_Treerings/Data/Adult_Field_Data_JAM_MCD.csv")
rings.diss <- Read_Tuscon("/Users/josh/Dropbox/Dissertation/CH1_Treerings/Data/Windendro_Dissertation")

#Match Tucson ringwidth files and adult census data for 50yrs of growth
combined.diss <- matchInventoryRings(trees=trees.diss,rings.diss)
combined.diss <- data.frame(combined.diss)
#write.csv(combined.diss, file="/Users/josh/Dropbox/Dissertation/CH1_Treerings/Data/combined.diss.new2.csv")
#manually added site average temp and precip
combined.diss.clim <- read.csv("/Users/josh/Dropbox/Dissertation/CH1_Treerings/Data/combined.diss.clim.csv")
#remove X before years
colnames.yrs <- c("1967","1968","1969","1970","1971","1972","1973","1974","1975","1976","1977","1978","1979","1980","1981","1982","1983","1984","1985","1986","1987","1988","1989","1990","1991","1992","1993","1994","1995","1996","1997","1998","1999","2000","2001","2002","2003","2004","2005","2006","2007","2008","2009","2010","2011","2012","2013","2014","2015","2016")

colnames(combined.diss.clim)[93:142] <- colnames.yrs#colnames(combined.dmg.clean) #get rid of stupid "x" before year

#TODO correct cores with known suspiciously large growth increment
#cores ID'd in /Users/josh/Dropbox/Dissertation/CH1_Treerings/Data/qc_removed_cores.xlsx

#write file to check which cores still need to be analyzed
#common sample ID to match with
combined.diss$sampleID <- mapply(combined.diss$SITE, combined.diss$PLOT, combined.diss$SUB, combined.diss$TAG, FUN = paste0)

trees.diss$sampleID <- mapply(trees.diss$Site, trees.diss$Plot, trees.diss$Sub, trees.diss$Tag, FUN = paste0)

analyzed.trees.growth <- merge(combined.diss, trees.diss, by = "sampleID", all.x = TRUE, all.y = TRUE)
check.growth.analyzed <- cbind(analyzed.trees.growth[,1:9],analyzed.trees.growth[,131:139],analyzed.trees.growth[,144:151],analyzed.trees.growth[,208],analyzed.trees.growth[,223])
names(check.growth.analyzed)[27] <- "DBH14"
names(check.growth.analyzed)[28] <- "DBH15"
write.csv(check.growth.analyzed, file = "alltrees_growth_analyzed_tracking_CDincluded.csv")

```
Reformat combined census and treering data dataframe to comply with format required for including time explicitly, which includes year by site and year by plot interaction terms.
```{r}
####didn't need to do all of this####
# #recreate data frame to be compliant with structure needed to include time as a variable (50 years)
# ids.rep.new.diss <- data.frame(matrix(NA,nrow=1,ncol=92))                    
# colnames(ids.rep.new.diss) <- colnames(combined.diss[1:92])
# for(i in 1:nrow(combined.diss)){
#   ids.rep.diss <- combined.diss[rep(i,times=50),1:92]
#   ids.rep.new.diss <- rbind(ids.rep.new.diss,ids.rep.diss)
# }
# #remove row of NAs
# combined.ids.diss <- ids.rep.new.diss[2:nrow(ids.rep.new.diss),]
# 
# #transpose and concatenate increments for all trees
# inc.new.diss <- vector("numeric",length=0)
# for(i in 1:nrow(combined.diss)){
#   inc.diss <- as.matrix(combined.diss[i,93:142])
#   inc.new.diss <- as.numeric(c(inc.new.diss,inc.diss))
# }
# 
# #add ring increments
# combined.ids.incs.diss <- cbind(combined.ids.diss,inc.new.diss)
# 
# #rename inc variable
# setnames(combined.ids.incs.diss,old='inc.new.diss',new='INCREMENT')
# 
# YEAR <- seq(1964,2013,1)
# YEAR <- rep(YEAR,1415)
# 
# #final version with 50yrs of growth
# combined.all.diss <- cbind(combined.ids.incs.diss,YEAR)
```

```{r}
#JAGS Bayesian Hierarchical linear model 
#Hierarchical terms for JAGS
ringwidths <- combined.diss.clim[97:136]
s<-as.numeric(as.factor(as.character(combined.diss.clim$SITE)))
p<-as.numeric(as.factor(paste0(as.character(combined.diss.clim$SITE),
                               as.character(combined.diss.clim$PLOT))))
sp <- as.numeric(as.factor(paste0(as.character(combined.diss.clim$SITE),
                                  as.character(combined.diss.clim$PLOT),
                                  as.character(combined.diss.clim$SUB))))
#creating interaction terms for JAGS
#Site by Year interaction, n = 400, 8 sites, 50 years each
#sy <- as.numeric(as.factor(paste0(as.character(combined.diss.clim$SITE),
                                  #colnames(combined.diss.clim[93:142]))))
# #Plot by Year interaction, n = 2600, 8 sites, 6-7 plots each
# spy <- as.numeric(as.factor(paste0(as.character(combined.diss.clim$SITE),
#                                   as.character(combined.diss.clim$PLOT),
#                                   colnames(combined.diss.clim[93:142]))))

#x = 1/var(combined.diss.clim$INCREMENT,na.rm=TRUE) #precision on all years
#y = mean(combined.diss.clim$INCREMENT,na.rm=TRUE) #mean all years
x = 1/var(as.vector(as.matrix(ringwidths)),na.rm=TRUE)
y = mean(as.matrix(ringwidths),na.rm=TRUE) ## overall mean growth rate
mean40yr = rowMeans(ringwidths,na.rm=TRUE) #means by individual
column.means = colMeans(ringwidths,na.rm=TRUE) #means by year
us = tapply(X=mean40yr,INDEX=list(s),FUN=mean,na.rm=TRUE)
up = tapply(X=mean40yr,INDEX=list(p),FUN=mean,na.rm=TRUE)
usp = tapply(X=mean40yr,INDEX=list(sp),FUN=mean,na.rm=TRUE)
#usy = tapply(X=mean50yr,INDEX=list(sy),FUN=mean,na.rm=TRUE)
usy = apply(ringwidths,2,function(x1){tapply(x1,s,FUN=mean,na.rm=TRUE)})
usy[is.nan(usy)] = 0
ty = 1/var(column.means)
ts = 1/var(us)
tp= 1/var(up)
tsp= 1/var(usp)
tsy= mean(1/var(usy))
as = us - mean(us)
ap = up - mean(up) 
asp= usp - mean(usp)
ai = mean40yr - mean(y)
ay = column.means - mean(column.means)

data2 = list(x=as.matrix(combined.diss.clim[,97:136]), nyear=40, 
             nrep = 1626,s=s,p=p,sp=sp,ns=max(s),np = max(p),nsp=max(sp))

asy = apply(data2$x,2,function(x){tapply(x,s,function(w){mean(w-y,na.rm=TRUE)})})
asy[is.nan(asy)]  = 0

init = list(mu = y,prec = x,tau.y = ty,tau.s = ts,tau.p = tp,tau.sp = tsp,
            tau.sy = tsy,tau.ind = x,alpha.site = as,alpha.plot = ap,
            alpha.sub = asp,alpha.siteyear = asy,alpha.ind = ai,alpha.y = ay)

FitRand.SiteYear = "
model {
mu ~ dnorm(0,0.001)
prec ~ dgamma(0.1,0.1)
tau.y ~ dgamma(0.1,0.1)  #prior year-effect
tau.s ~ dgamma(0.1,0.1)  #prior site
tau.p ~ dgamma(0.1,0.1)  #prior plot
tau.sp ~ dgamma(0.1,0.1) #prior subplot
tau.ind ~ dgamma(0.1,0.1) #prior individual
tau.sy ~ dgamma(0.1,0.1) #prior site x year effect
#tau.spy ~ dgamma(0.1,0.1) #prior plot x year effect

for(s in 1:ns){
  alpha.site[s] ~ dnorm(0,tau.s)
  for(y in 1:nyear){
    alpha.siteyear[s,y] ~ dnorm(0,tau.sy)
  }
}
for(p in 1:np){
alpha.plot[p] ~ dnorm(0,tau.p)
}
for(sp in 1:nsp){
alpha.sub[sp] ~ dnorm(0,tau.sp)
}
for(i in 1:nrep){
alpha.ind[i] ~ dnorm(0,tau.ind)
}

for(y in 1:nyear){
alpha.y[y] ~dnorm(0,tau.y)
for(i in 1:nrep){
Ex[i,y] <- mu + alpha.y[y]  + alpha.site[s[i]] + alpha.plot[p[i]] + alpha.sub[sp[i]] + alpha.ind[i] + alpha.siteyear[s[i],y]

x[i,y] ~dnorm(Ex[i,y],prec)
}
}
}"

## compile JAGS model
j.model   <- jags.model (file = textConnection(FitRand.SiteYear),
                         data = data2,
                         inits = init,
                         n.chains = 3)
## burn-in
b   <- coda.samples (model = j.model,
                      variable.names = c("mu","prec","tau.y","tau.s","tau.p","tau.sp","tau.ind","tau.sy"),
                      n.iter = 1000,
                      thin = 2)

# bt   <- coda.samples (model = j.model,
#                      variable.names = c("mu","prec","tau.y","tau.s","tau.p","tau.sp","tau.ind"),
#                      n.iter = 5000)

update.jags<-function(model,previous.coda,variable.names,iter){
  samp <- coda.samples (model = model,
                        variable.names = variable.names,
                        n.iter = iter,
                        thin=2)
  new.list <- list()
  for(i in 1:length(samp)){
    new.list[[i]] = as.mcmc(rbind(previous.coda[[i]],samp[[i]]))
  }
  bmcmc <<- mcmc.list(new.list)
  #return(bmcmc)
}

update.jags(j.model,b,c("mu","prec","tau.y","tau.s","tau.p","tau.sp","tau.ind","tau.sy"),500)
b.out<-b
bmcmc1<-bmcmc

update.jags(j.model,bmcmc1,c("mu","prec","tau.y","tau.s","tau.p","tau.sp","tau.ind","tau.sy"),1000)
bmcmc2<-bmcmc

update.jags(j.model,bmcmc2,c("mu","prec","tau.y","tau.s","tau.p","tau.sp","tau.ind","tau.sy"),2000)
bmcmc3<-bmcmc

update.jags(j.model,bmcmc3,c("mu","prec","tau.y","tau.s","tau.p","tau.sp","tau.ind","tau.sy"),2000)
bmcmc4<-bmcmc

update.jags(j.model,bmcmc4,c("mu","prec","tau.y","tau.s","tau.p","tau.sp","tau.ind","tau.sy"),5000)
bmcmc5<-bmcmc

update.jags(j.model,bmcmc5,c("mu","prec","tau.y","tau.s","tau.p","tau.sp","tau.ind","tau.sy"),5000)
bmcmc6<-bmcmc

# update.jags(j.model,bmcmc6,c("mu","prec","tau.y","tau.s","tau.p","tau.sp","tau.ind"),20000)
# bmcmc7<-bmcmc
# update.jags(j.model,bmcmc7,c("mu","prec","tau.y","tau.s","tau.p","tau.sp","tau.ind"),10000)
# bmcmc8<-bmcmc
# update.jags(j.model,bmcmc8,c("mu","prec","tau.y","tau.s","tau.p","tau.sp","tau.ind"),20000)
# bmcmc9<-bmcmc

#transform posteriors to standard deviations
test.sd2 <- list()
test.sd2 <- for(i in 1:3){
  test.sd2[[i]]<-as.mcmc(sqrt(1/bmcmc1[[i]]))

  test.SD <- mcmc.list(test.sd2)
}

for(i in 1:8){
  plot(test.SD[,i],cex.lab=1.8,xlim=c(0,1.25))
}


str(summary(bmcmc1))
#for(i in 1:7){
  summary(bmcmc1)#[,i])    ## summary table
#}

for(i in 1:8){
  plot(b.out[,i])
  #plot(sqrt(1/(bmcmc9[,i]))) ## mcmc trace and density plots
}

# for(i in 1:7){
#   plot(bmcmc7[,i]) ## mcmc trace and density plots
# }
# 
# for(i in 1:7){
#   plot(bmcmc6[,i]) ## mcmc trace and density plots
# }
# 
# for(i in 1:7){
#   plot(bmcmc5[,i]) ## mcmc trace and density plots
# }
# for(i in 1:7){
#   plot(bmcmc3[,i]) ## mcmc trace and density plots
# }
# #bmcmc is new output
# for(i in 1:7){
#   plot(bmcmc2[,i]) ## mcmc trace and density plots
# }
# for(i in 1:7){
#   plot(b.out[,i]) ## mcmc trace and density plots
# }
for(i in 1:8){
autocorr.plot(b.out[,i]) ## autocorrelation
}
for(i in 1:8){
cumuplot(b.out[,i]) ## quantile plot
}
for(i in 1:8){
gelman.plot(b.out[,i])      ## GRB statistic
}
for(i in 1:8){
summary(b.out[,i])    ## summary table
}

  ####plot of all individuals####
denplot(test.SD[,2:7],collapse=TRUE,xlim=c(0,1.25),main=c("Residual",
                                                          "Individual",
                                                          "Landscape",
                                                          "Region",
                                                          "Microsite",
                                                          "Year"))

##summary stats for making barplot of median variance explained
summary.all <- summary(test.SD) 
var.all<-summary.all$quantiles[2:8,3]
#proportion of variance explained
proportion.var.all<- var.all/sum(var.all)

##### done visualizing outputs #####

update.jags<-function(model,previous.coda,variable.names,iter){
  samp <- coda.samples (model = model,
                variable.names = variable.names,
                n.iter = iter)
  bmcmc <<- mcmc(do.call(rbind,c(previous.coda,samp)))
  #return(bmcmc)
}

update.jags(j.model,b,c("mu","prec","tau.y","tau.s","tau.p","tau.sp","tau.ind"),100)
# b3 <- coda.samples (model = j.model,
#                     variable.names = c("mu","prec"),#,"alpha.y","alpha.site","alpha.plot","alpha.sub"),#,"alpha.ind"),
#                     n.iter = 500)#,
# update.jags(model=j.model,previous.coda=b3,variable.names=c("mu","prec"),iter=500)

# bmcmc3 = b3
# plot(bmcmc3)



mu2 = as.data.frame(as.matrix(bmcmc2))$mu
quantile(mu2,c(0.025,0.5,0.975))







# #just look at AGU code for example of how to make unique IDs
# #Example creating interaction term
# Y = seq(1,10)
# L = as.vector(x=c("A","B","C"))
# YL = expand.grid(Y,L) #creates matrix of all combinations as characters
# YL.vector = as.vector(paste0(YL$Var1,YL$Var2)) #vector of all combinations as characters
# #Can now use this to create interaction term for years x plot and years x site

```

```{r}
#Add covariates to data

count(combined.diss.clim$SPP)
#remove spp with very low sample size, setting cutoff below 5 because
#of an arbitrary reason, need to do an actual power analysis

abba <- which(combined.diss.clim$SPP == "ABBA")
#acne2 <- which(combined.diss.clim$SPP == "ACNE2")
acru <- which(combined.diss.clim$SPP == "ACRU")
acsa3 <- which(combined.diss.clim$SPP == "ACSA3")
beal2 <- which(combined.diss.clim$SPP == "BEAL2")
bele <- which(combined.diss.clim$SPP == "BELE")
bepa <- which(combined.diss.clim$SPP == "BEPA")
#caco15 <- which(combined.diss.clim$SPP == "CACO15")
cagl8 <- which(combined.diss.clim$SPP == "CAGL8")
#caov2 <- which(combined.diss.clim$SPP == "CAOV2")
cato6 <- which(combined.diss.clim$SPP == "CATO6")
fagr <- which(combined.diss.clim$SPP == "FAGR")
fram2 <- which(combined.diss.clim$SPP == "FRAM2")
frni <- which(combined.diss.clim$SPP == "FRNI")
#juni <- which(combined.diss.clim$SPP == "JUNI")
juvi <- which(combined.diss.clim$SPP == "JUVI")
#list2 <- which(combined.diss.clim$SPP == "LIST2")
litu <- which(combined.diss.clim$SPP == "LITU")
mavi2 <- which(combined.diss.clim$SPP == "MAVI2")
nysy <- which(combined.diss.clim$SPP == "NYSY")
oxar <- which(combined.diss.clim$SPP == "OXAR")
pima <- which(combined.diss.clim$SPP == "PIMA")
pipa2 <- which(combined.diss.clim$SPP == "PIPA2")
#piri <- which(combined.diss.clim$SPP == "PIRI")
piru <- which(combined.diss.clim$SPP == "PIRU")
pist <- which(combined.diss.clim$SPP == "PIST")
pita <- which(combined.diss.clim$SPP == "PITA")
pogr4 <- which(combined.diss.clim$SPP == "POGR4")
potr5 <- which(combined.diss.clim$SPP == "POTR5")
prse2 <- which(combined.diss.clim$SPP == "PRSE2")
qual <- which(combined.diss.clim$SPP == "QUAL")
quco2 <- which(combined.diss.clim$SPP == "QUCO2")
qula2 <- which(combined.diss.clim$SPP == "QULA2")
qula3 <- which(combined.diss.clim$SPP == "QULA3")
qumo4 <- which(combined.diss.clim$SPP == "QUMO4")
qumu <- which(combined.diss.clim$SPP == "QUMU")
quni <- which(combined.diss.clim$SPP == "QUNI")
#quph <- which(combined.diss.clim$SPP == "QUPH")
quru <- which(combined.diss.clim$SPP == "QURU")
quve <- which(combined.diss.clim$SPP == "QUVE")
#quvi <- which(combined.diss.clim$SPP == "QUVI")
saal5 <- which(combined.diss.clim$SPP == "SAAL5")
thoc2 <- which(combined.diss.clim$SPP == "THOC2")
#tiam <- which(combined.diss.clim$SPP == "TIAM")
tsca <- which(combined.diss.clim$SPP == "TSCA")
ulam <- which(combined.diss.clim$SPP == "ULAM")

combined.diss.clim.common <- rbind(
                             combined.diss.clim[abba,],
                             combined.diss.clim[acru,],
                             combined.diss.clim[acsa3,],
                             combined.diss.clim[beal2,],
                             combined.diss.clim[bele,],
                             combined.diss.clim[bepa,],
                             combined.diss.clim[cagl8,],
                             combined.diss.clim[cato6,],
                             combined.diss.clim[fagr,],
                             combined.diss.clim[fram2,],
                             combined.diss.clim[frni,],
                             combined.diss.clim[juvi,],
                             combined.diss.clim[litu,],
                             combined.diss.clim[mavi2,],
                             combined.diss.clim[nysy,],
                             combined.diss.clim[oxar,],
                             combined.diss.clim[pima,],
                             combined.diss.clim[pipa2,],
                             combined.diss.clim[piru,],
                             combined.diss.clim[pist,],
                             combined.diss.clim[pita,],
                             combined.diss.clim[pogr4,],
                             combined.diss.clim[potr5,],
                             combined.diss.clim[prse2,],
                             combined.diss.clim[quco2,],
                             combined.diss.clim[qual,],
                             combined.diss.clim[qula2,],
                             combined.diss.clim[qula3,],
                             combined.diss.clim[qumo4,],
                             combined.diss.clim[qumu,],
                             combined.diss.clim[quni,],
                             combined.diss.clim[quru,],
                             combined.diss.clim[quve,],
                             combined.diss.clim[saal5,],
                             combined.diss.clim[thoc2,],
                             combined.diss.clim[tsca,],
                             combined.diss.clim[ulam,])
### Adding PFT and Wood Anatomy covariates ###

##adding PFTs
###Reference: http://plants.usda.gov/java/
### Also USFS silvics manual: https://www.na.fs.fed.us/spfo/pubs/silvics_manual/table_of_contents.htm

# Hardwoods
Early.Hardwood <- c(
                    which(combined.diss.clim.common$SPP == "ACPE"),#
                    which(combined.diss.clim.common$SPP == "AIAL"),#
                    which(combined.diss.clim.common$SPP == "BEAL2"),#
                    which(combined.diss.clim.common$SPP == "BELE"),#
                    which(combined.diss.clim.common$SPP == "BEPA"),#
                    which(combined.diss.clim.common$SPP == "BETUL"),#
                    which(combined.diss.clim.common$SPP == "COFL2"),#
                    which(combined.diss.clim.common$SPP == "JUCI"),#
                    which(combined.diss.clim.common$SPP == "JUNI"),#
                    which(combined.diss.clim.common$SPP == "LIST2"),#
                    which(combined.diss.clim.common$SPP == "LITU"),#
                    which(combined.diss.clim.common$SPP == "POGR4"),#
                    which(combined.diss.clim.common$SPP == "POTR5"),#
                    which(combined.diss.clim.common$SPP == "PRSE2"),#
                    which(combined.diss.clim.common$SPP == "ROPS"),#
                    which(combined.diss.clim.common$SPP == "SAAL5")#
                    ) #16 Spp

Mid.Hardwood <- c(
                  which(combined.diss.clim.common$SPP == "FRAM2"),#
                  which(combined.diss.clim.common$SPP == "FRNI"),#
                  which(combined.diss.clim.common$SPP == "FRQU"),#
                  which(combined.diss.clim.common$SPP == "ACNE2"),#
                  which(combined.diss.clim.common$SPP == "ACRU"),#
                  which(combined.diss.clim.common$SPP == "CACA18"),#
                  which(combined.diss.clim.common$SPP == "CACO15"),#
                  which(combined.diss.clim.common$SPP == "CAGL8"),#
                  which(combined.diss.clim.common$SPP == "CALA21"),#
                  which(combined.diss.clim.common$SPP == "CAOV2"),#
                  which(combined.diss.clim.common$SPP == "CATO6"),#
                  which(combined.diss.clim.common$SPP == "CECA4"),#
                  which(combined.diss.clim.common$SPP == "MAAC"),#
                  which(combined.diss.clim.common$SPP == "MAFR"),#
                  which(combined.diss.clim.common$SPP == "MAVI2"),#
                  which(combined.diss.clim.common$SPP == "PLOC"),#
                  which(combined.diss.clim.common$SPP == "QUAL"),#
                  which(combined.diss.clim.common$SPP == "QUCO2"),#
                  which(combined.diss.clim.common$SPP == "QUERC"),#
                  which(combined.diss.clim.common$SPP == "QUIM"),#
                  which(combined.diss.clim.common$SPP == "QULA2"),#
                  which(combined.diss.clim.common$SPP == "QULA3"),#
                  which(combined.diss.clim.common$SPP == "QUMO4"),#
                  which(combined.diss.clim.common$SPP == "QUMU"),#
                  which(combined.diss.clim.common$SPP == "QUNI"),#
                  which(combined.diss.clim.common$SPP == "QUPH"),#
                  which(combined.diss.clim.common$SPP == "QURU"),#
                  which(combined.diss.clim.common$SPP == "QUST"),#
                  which(combined.diss.clim.common$SPP == "QUVE"),#
                  which(combined.diss.clim.common$SPP == "QUVI"),#
                  which(combined.diss.clim.common$SPP == "ULAM"),
                  which(combined.diss.clim.common$SPP == "ULAL")#
                  ) #32 Spp

Late.Hardwood <- c(
                  which(combined.diss.clim.common$SPP == "ACFL"),#
                  which(combined.diss.clim.common$SPP == "ACSA3"),#
                  which(combined.diss.clim.common$SPP == "AEGL"),#
                  which(combined.diss.clim.common$SPP == "DIVI5"),#
                  which(combined.diss.clim.common$SPP == "FAGR"),#
                  which(combined.diss.clim.common$SPP == "MAGR4"),#
                  which(combined.diss.clim.common$SPP == "NYSY"),#
                  which(combined.diss.clim.common$SPP == "OXAR"),#
                  which(combined.diss.clim.common$SPP == "TIAM")#
                  ) #9 Spp

# set PFTs
combined.diss.clim.common$PFT <- NA
combined.diss.clim.common$PFT[Early.Hardwood] <- "Early.Hardwood"
combined.diss.clim.common$PFT[Mid.Hardwood] <- "Mid.Hardwood"
combined.diss.clim.common$PFT[Late.Hardwood] <- "Late.Hardwood"

# Conifers
Early.Conifer <- c(
                  which(combined.diss.clim.common$SPP == "ABBA"),#
                  which(combined.diss.clim.common$SPP == "JUVI"),#
                  which(combined.diss.clim.common$SPP == "PIPA2"),#
                  which(combined.diss.clim.common$SPP == "PIST"),#
                  which(combined.diss.clim.common$SPP == "PIVI2"),#
                  which(combined.diss.clim.common$SPP == "PIRE"),#
                  which(combined.diss.clim.common$SPP == "PIRI"),#
                  which(combined.diss.clim.common$SPP == "PITA")#
                  ) #8 Spp

Late.Conifer <- c(
                  which(combined.diss.clim.common$SPP == "LALA"),#
                  which(combined.diss.clim.common$SPP == "PIGL"),#
                  which(combined.diss.clim.common$SPP == "PIMA"),#
                  which(combined.diss.clim.common$SPP == "PIRU"),#
                  which(combined.diss.clim.common$SPP == "THOC2"),#
                  which(combined.diss.clim.common$SPP == "TSCA")#
                  ) #6 Spp

combined.diss.clim.common$PFT[Early.Conifer] <- "Early.Conifer"
combined.diss.clim.common$PFT[Late.Conifer] <- "Late.Conifer"

# adding wood anatomy
###Reference http://www.wood-database.com/
###Disagree with reference for hickory classification - put all as semi-ring porous. conferred with Christy

large.resin <- c(
                which(combined.diss.clim.common$SPP == "PIPA2"),
                which(combined.diss.clim.common$SPP == "PIST"),
                which(combined.diss.clim.common$SPP == "PIVI2"),
                which(combined.diss.clim.common$SPP == "PIRE"),
                which(combined.diss.clim.common$SPP == "PIRI"),
                which(combined.diss.clim.common$SPP == "PITA")
                )

small.resin <- c(
                which(combined.diss.clim.common$SPP == "LALA"),
                which(combined.diss.clim.common$SPP == "PIGL"),
                which(combined.diss.clim.common$SPP == "PIMA"),
                which(combined.diss.clim.common$SPP == "PIRU")
                )

no.resin <- c(
              which(combined.diss.clim.common$SPP == "ABBA"),
              which(combined.diss.clim.common$SPP == "JUVI"),
              which(combined.diss.clim.common$SPP == "THOC2"),
              which(combined.diss.clim.common$SPP == "TSCA")
              )

ring.porous <- c(
                which(combined.diss.clim.common$SPP == "AIAL"),              
                which(combined.diss.clim.common$SPP == "FRAM2"),
                which(combined.diss.clim.common$SPP == "FRNI"),
                which(combined.diss.clim.common$SPP == "FRQU"),
                which(combined.diss.clim.common$SPP == "QUAL"),
                which(combined.diss.clim.common$SPP == "QUCO2"),
                which(combined.diss.clim.common$SPP == "QUERC"),
                which(combined.diss.clim.common$SPP == "QUIM"),
                which(combined.diss.clim.common$SPP == "QULA2"),
                which(combined.diss.clim.common$SPP == "QULA3"),
                which(combined.diss.clim.common$SPP == "QUMO4"),
                which(combined.diss.clim.common$SPP == "QUMU"),
                which(combined.diss.clim.common$SPP == "QUNI"),
                which(combined.diss.clim.common$SPP == "QUPH"),
                which(combined.diss.clim.common$SPP == "QURU"),
                which(combined.diss.clim.common$SPP == "QUST"),
                which(combined.diss.clim.common$SPP == "QUVE"),
                which(combined.diss.clim.common$SPP == "QUVI"),
                which(combined.diss.clim.common$SPP == "ROPS"),
                which(combined.diss.clim.common$SPP == "SAAL5")
                )

diffuse.porous <- c(
                    which(combined.diss.clim.common$SPP == "ACNE2"),
                    which(combined.diss.clim.common$SPP == "ACPE"),
                    which(combined.diss.clim.common$SPP == "ACRU"),
                    which(combined.diss.clim.common$SPP == "ACSA3"),
                    which(combined.diss.clim.common$SPP == "AEGL"),
                    which(combined.diss.clim.common$SPP == "BEAL2"),
                    which(combined.diss.clim.common$SPP == "BELE"),
                    which(combined.diss.clim.common$SPP == "BEPA"),
                    which(combined.diss.clim.common$SPP == "BETUL"),
                    which(combined.diss.clim.common$SPP == "CACA18"),
                    which(combined.diss.clim.common$SPP == "COFL2"),
                    which(combined.diss.clim.common$SPP == "LIST2"),
                    which(combined.diss.clim.common$SPP == "LITU"),
                    which(combined.diss.clim.common$SPP == "MAAC"),
                    which(combined.diss.clim.common$SPP == "MAFR"),
                    which(combined.diss.clim.common$SPP == "MAVI2"),                  
                    which(combined.diss.clim.common$SPP == "PLOC"),
                    which(combined.diss.clim.common$SPP == "FAGR"),
                    which(combined.diss.clim.common$SPP == "MAGR4"),
                    which(combined.diss.clim.common$SPP == "NYSY"),
                    which(combined.diss.clim.common$SPP == "OXAR"),
                    which(combined.diss.clim.common$SPP == "TIAM"),
                    which(combined.diss.clim.common$SPP == "POGR4"),
                    which(combined.diss.clim.common$SPP == "POTR5"),
                    which(combined.diss.clim.common$SPP == "PRSE2")               
                    )

semiring.porous <- c(
                    which(combined.diss.clim.common$SPP == "CACO15"),
                    which(combined.diss.clim.common$SPP == "CAGL8"),
                    which(combined.diss.clim.common$SPP == "CALA21"),
                    which(combined.diss.clim.common$SPP == "CAOV2"),
                    which(combined.diss.clim.common$SPP == "CATO6"),
                    which(combined.diss.clim.common$SPP == "CECA4"),
                    which(combined.diss.clim.common$SPP == "DIVI5"), #bordering diffuse porous
                    which(combined.diss.clim.common$SPP == "JUCI"),
                    which(combined.diss.clim.common$SPP == "JUNI"),
                    which(combined.diss.clim.common$SPP == "ULAM"),
                    which(combined.diss.clim.common$SPP == "ULAL")
                    )

#set anatomy
combined.diss.clim.common$WOOD <- NA
combined.diss.clim.common$WOOD[large.resin] <- "large.resin"
combined.diss.clim.common$WOOD[small.resin] <- "small.resin"
combined.diss.clim.common$WOOD[no.resin] <- "no.resin"
combined.diss.clim.common$WOOD[ring.porous] <- "ring.porous"
combined.diss.clim.common$WOOD[diffuse.porous] <- "diffuse.porous"
combined.diss.clim.common$WOOD[semiring.porous] <- "semiring.porous"

#Relevel Factors
combined.diss.clim.common$WOOD <- factor(combined.diss.clim.common$WOOD, 
                                    levels = c("large.resin","small.resin","no.resin",
                                               "ring.porous","semiring.porous",
                                               "diffuse.porous"), ordered = TRUE)

#write file to see if missing any species
write.csv(combined.diss.clim.common, file="/Users/josh/Dropbox/Dissertation/CH1_Treerings/Data/combined.diss.clim.common.wood.csv")

################## Haven't added species and other phylogeny yet 3/30/2017 ####################
###############################################################################################
###Reference: http://plants.usda.gov/java/
# adding species string
    alba <- which(combined.diss.clim.common$SPP == "QUAL")
    coccinea <- which(combined.diss.clim.common$SPP == "QUCO2")
    none <- c(which(combined.diss.clim.common$SPP == "QUERC"),
              which(combined.diss.clim.common$SPP == "BETUL"))
    imbricaria <- which(combined.diss.clim.common$SPP == "QUIM")
    laevis <- which(combined.diss.clim.common$SPP == "QULA2")
    laurifolia <- which(combined.diss.clim.common$SPP == "QULA3")
    montana <- which(combined.diss.clim.common$SPP == "QUMO4")
    muehlenbergii <- which(combined.diss.clim.common$SPP == "QUMU")
    nigra <- c(which(combined.diss.clim.common$SPP == "QUNI"),
               which(combined.diss.clim.common$SPP == "FRNI"),
               which(combined.diss.clim.common$SPP == "JUNI"))
    phellos <- which(combined.diss.clim.common$SPP == "QUPH")
    rubra <- which(combined.diss.clim.common$SPP == "QURU")
    stellata <- which(combined.diss.clim.common$SPP == "QUST")
    velutina <- which(combined.diss.clim.common$SPP == "QUVE")
    virginiana <- c(which(combined.diss.clim.common$SPP == "QUVI"),
                  which(combined.diss.clim.common$SPP == "MAVI2"),
                  which(combined.diss.clim.common$SPP == "DIVI5"),
                  which(combined.diss.clim.common$SPP == "PIVI2"),
                  which(combined.diss.clim.common$SPP == "JUVI"))
    americana <- c(which(combined.diss.clim.common$SPP == "FRAM2"),
                   which(combined.diss.clim.common$SPP == "TIAM"),
                   which(combined.diss.clim.common$SPP == "ULAM"))
    quadrangulata <- which(combined.diss.clim.common$SPP == "FRQU")
    altissima <- which(combined.diss.clim.common$SPP == "AIAL")
    pseudoacacia <- which(combined.diss.clim.common$SPP == "ROPS")
    albidum <- which(combined.diss.clim.common$SPP == "SAAL5")
    negundo <- which(combined.diss.clim.common$SPP == "ACNE2")
    pensylvanicum <- which(combined.diss.clim.common$SPP == "ACPE")
    rubrum <- which(combined.diss.clim.common$SPP == "ACRU")
    saccharum <- which(combined.diss.clim.common$SPP == "ACSA3")
    glabra <- c(which(combined.diss.clim.common$SPP == "AEGL"),
                which(combined.diss.clim.common$SPP == "CAGL8"))
    alleghaniensis <- which(combined.diss.clim.common$SPP == "BEAL2")
    lenta <- which(combined.diss.clim.common$SPP == "BELE")
    papyrifera <- which(combined.diss.clim.common$SPP == "BEPA")
    caroliniana <- which(combined.diss.clim.common$SPP == "CACA18")
    florida <- which(combined.diss.clim.common$SPP == "COFL2")
    styraciflua <- which(combined.diss.clim.common$SPP == "LIST2")
    tulipifera <- which(combined.diss.clim.common$SPP == "LITU")
    acuminata <- which(combined.diss.clim.common$SPP == "MAAC")
    fraseri <- which(combined.diss.clim.common$SPP == "MAFR")
    grandiflora <- which(combined.diss.clim.common$SPP == "MAGR4")
    occidentalis <- which(combined.diss.clim.common$SPP == "PLOC")
    grandifolia <- which(combined.diss.clim.common$SPP == "FAGR")
    sylvatica <- which(combined.diss.clim.common$SPP == "NYSY")
    arboreum <- which(combined.diss.clim.common$SPP == "OXAR")
    grandidentata <- which(combined.diss.clim.common$SPP == "POGR4")
    tremuloides <- which(combined.diss.clim.common$SPP == "POTR5")
    serotina <- which(combined.diss.clim.common$SPP == "PRSE2")
    cordiformis <- which(combined.diss.clim.common$SPP == "CACO15")
    laciniosa <- which(combined.diss.clim.common$SPP == "CALA21")
    ovata <- which(combined.diss.clim.common$SPP == "CAOV2")
    tomentosa <- which(combined.diss.clim.common$SPP == "CATO6")
    canadensis <- c(which(combined.diss.clim.common$SPP == "CECA4"),
                    which(combined.diss.clim.common$SPP == "TSCA"))
    cinerea <- which(combined.diss.clim.common$SPP == "JUCI")
    palustris <- which(combined.diss.clim.common$SPP == "PIPA2")
    strobus <- which(combined.diss.clim.common$SPP == "PIST")
    resinosa <- which(combined.diss.clim.common$SPP == "PIRE")
    rigida <- which(combined.diss.clim.common$SPP == "PIRI")
    taeda <- which(combined.diss.clim.common$SPP == "PITA")
    laricina <- which(combined.diss.clim.common$SPP == "LALA")
    glauca <- which(combined.diss.clim.common$SPP == "PIGL")
    mariana <- which(combined.diss.clim.common$SPP == "PIMA")
    rubens <- which(combined.diss.clim.common$SPP == "PIRU")
    balsamea <- which(combined.diss.clim.common$SPP == "ABBA")
    occidentalis <- which(combined.diss.clim.common$SPP == "THOC2")
    
    ### Set total dataframe for species
    
    combined.diss.clim.common$species <- NA
    combined.diss.clim.common$species[alba] <- "alba"
    combined.diss.clim.common$species[coccinea] <- "coccinea"
    combined.diss.clim.common$species[none] <- "none"
    combined.diss.clim.common$species[imbricaria] <- "imbricaria"
    combined.diss.clim.common$species[laevis] <- "laevis"
    combined.diss.clim.common$species[laurifolia] <- "laurifolia"
    combined.diss.clim.common$species[montana] <- "montana"
    combined.diss.clim.common$species[muehlenbergii] <- "muehlenbergii"
    combined.diss.clim.common$species[nigra] <- "nigra"
    combined.diss.clim.common$species[phellos] <- "phellos"
    combined.diss.clim.common$species[rubra] <- "rubra"
    combined.diss.clim.common$species[stellata] <- "stellata"
    combined.diss.clim.common$species[velutina] <- "velutina"
    combined.diss.clim.common$species[virginiana] <- "virginiana"
    combined.diss.clim.common$species[americana] <- "americana"
    combined.diss.clim.common$species[quadrangulata] <- "quadrangulata"
    combined.diss.clim.common$species[altissima] <- "altissima"
    combined.diss.clim.common$species[pseudoacacia] <- "pseudoacacia"
    combined.diss.clim.common$species[albidum] <- "albidum"
    combined.diss.clim.common$species[negundo] <- "negundo"
    combined.diss.clim.common$species[pensylvanicum] <- "pensylvanicum"
    combined.diss.clim.common$species[rubrum] <- "rubrum"
    combined.diss.clim.common$species[saccharum] <- "saccharum"
    combined.diss.clim.common$species[glabra] <- "glabra"
    combined.diss.clim.common$species[alleghaniensis] <- "alleghaniensis"
    combined.diss.clim.common$species[lenta] <- "lenta"
    combined.diss.clim.common$species[papyrifera] <- "papyrifera"
    combined.diss.clim.common$species[caroliniana] <- "caroliniana"
    combined.diss.clim.common$species[florida] <- "florida"
    combined.diss.clim.common$species[styraciflua] <- "styraciflua"
    combined.diss.clim.common$species[tulipifera] <- "tulipifera"
    combined.diss.clim.common$species[acuminata] <- "acuminata"
    combined.diss.clim.common$species[fraseri] <- "fraseri"
    combined.diss.clim.common$species[grandiflora] <- "grandiflora"
    combined.diss.clim.common$species[occidentalis] <- "occidentalis"
    combined.diss.clim.common$species[grandifolia] <- "grandifolia"
    combined.diss.clim.common$species[sylvatica] <- "sylvatica"
    combined.diss.clim.common$species[arboreum] <- "arboreum"
    combined.diss.clim.common$species[grandidentata] <- "grandidentata"
    combined.diss.clim.common$species[tremuloides] <- "tremuloides"
    combined.diss.clim.common$species[serotina] <- "serotina"
    combined.diss.clim.common$species[cordiformis] <- "cordiformis"
    combined.diss.clim.common$species[laciniosa] <- "laciniosa"
    combined.diss.clim.common$species[ovata] <- "ovata"
    combined.diss.clim.common$species[tomentosa] <- "tomentosa"
    combined.diss.clim.common$species[canadensis] <- "canadensis"
    combined.diss.clim.common$species[cinerea] <- "cinerea"
    combined.diss.clim.common$species[palustris] <- "palustris"
    combined.diss.clim.common$species[strobus] <- "strobus"
    combined.diss.clim.common$species[resinosa] <- "resinosa"
    combined.diss.clim.common$species[rigida] <- "rigida"
    combined.diss.clim.common$species[taeda] <- "taeda"
    combined.diss.clim.common$species[laricina] <- "laricina"
    combined.diss.clim.common$species[glauca] <- "glauca"
    combined.diss.clim.common$species[mariana] <- "mariana"
    combined.diss.clim.common$species[rubens] <- "rubens"
    combined.diss.clim.common$species[balsamea] <- "balsamea"
    combined.diss.clim.common$species[occidentalis] <- "occidentalis"

# adding genus and family
  ### Hardwoods

    Quercus <- c(                
                which(combined.diss.clim.common$SPP == "QUAL"),
                which(combined.diss.clim.common$SPP == "QUCO2"),
                which(combined.diss.clim.common$SPP == "QUERC"),
                which(combined.diss.clim.common$SPP == "QUIM"),
                which(combined.diss.clim.common$SPP == "QULA2"),
                which(combined.diss.clim.common$SPP == "QULA3"),
                which(combined.diss.clim.common$SPP == "QUMO4"),
                which(combined.diss.clim.common$SPP == "QUMU"),
                which(combined.diss.clim.common$SPP == "QUNI"),
                which(combined.diss.clim.common$SPP == "QUPH"),
                which(combined.diss.clim.common$SPP == "QURU"),
                which(combined.diss.clim.common$SPP == "QUST"),
                which(combined.diss.clim.common$SPP == "QUVE"),
                which(combined.diss.clim.common$SPP == "QUVI")
                )
    
   Fraxinus <- c(
                which(combined.diss.clim.common$SPP == "FRAM2"),
                which(combined.diss.clim.common$SPP == "FRNI"),
                which(combined.diss.clim.common$SPP == "FRQU")
                )
   
  Ailanthus <- c(
                which(combined.diss.clim.common$SPP == "AIAL")
                )
   
    Robinia <- c(
                which(combined.diss.clim.common$SPP == "ROPS")
                )
    
  Sassafras <- c(
                which(combined.diss.clim.common$SPP == "SAAL5")
                )
  
      Acer <- c(
                which(combined.diss.clim.common$SPP == "ACNE2"),
                which(combined.diss.clim.common$SPP == "ACPE"),
                which(combined.diss.clim.common$SPP == "ACRU"),
                which(combined.diss.clim.common$SPP == "ACSA3")
                )
      
  Aesculus <- c(
                which(combined.diss.clim.common$SPP == "AEGL")
                )
  
    Betula <- c(
                which(combined.diss.clim.common$SPP == "BEAL2"),
                which(combined.diss.clim.common$SPP == "BELE"),
                which(combined.diss.clim.common$SPP == "BEPA"),
                which(combined.diss.clim.common$SPP == "BETUL")
                )
    
  Carpinus <- c(
                which(combined.diss.clim.common$SPP == "CACA18")
                )
    
    Cornus <- c(
                which(combined.diss.clim.common$SPP == "COFL2")
                )
    
Liquidambar <- c(
                which(combined.diss.clim.common$SPP == "LIST2")
                )

Liriodendron <- c(
                which(combined.diss.clim.common$SPP == "LITU")
                )

    Magnolia <- c(
                which(combined.diss.clim.common$SPP == "MAAC"),
                which(combined.diss.clim.common$SPP == "MAFR"),
                which(combined.diss.clim.common$SPP == "MAVI2"),
                which(combined.diss.clim.common$SPP == "MAGR4")
                )
    
    Platanus <- c(
                which(combined.diss.clim.common$SPP == "PLOC")
                )
    
       Fagus <- c(
                which(combined.diss.clim.common$SPP == "FAGR")
                )
       
       Nyssa <- c(
                which(combined.diss.clim.common$SPP == "NYSY")
                )
       
       Oxydendrum <- c(
                which(combined.diss.clim.common$SPP == "OXAR")
                )
       
       Tilia <- c(
                which(combined.diss.clim.common$SPP == "TIAM")
                )
    
     Populus <- c(
                which(combined.diss.clim.common$SPP == "POGR4"),
                which(combined.diss.clim.common$SPP == "POTR5")
                )
       
     Prunus <- c(
                which(combined.diss.clim.common$SPP == "PRSE2")
                )

       Carya <- c(
                which(combined.diss.clim.common$SPP == "CACO15"),
                which(combined.diss.clim.common$SPP == "CAGL8"),
                which(combined.diss.clim.common$SPP == "CALA21"),
                which(combined.diss.clim.common$SPP == "CAOV2"),
                which(combined.diss.clim.common$SPP == "CATO6")
                )
       
     Cercis <- c(
                which(combined.diss.clim.common$SPP == "CECA4")
                )
     
  Diospyros <- c(
                which(combined.diss.clim.common$SPP == "DIVI5")
                )
  
    Juglans <- c(
                which(combined.diss.clim.common$SPP == "JUCI"),
                which(combined.diss.clim.common$SPP == "JUNI")
                )
    
      Ulmus <- c(
                which(combined.diss.clim.common$SPP == "ULAM")
                )

  ### Conifers
      Pinus <- c(
                which(combined.diss.clim.common$SPP == "PIPA2"),
                which(combined.diss.clim.common$SPP == "PIST"),
                which(combined.diss.clim.common$SPP == "PIVI2"),
                which(combined.diss.clim.common$SPP == "PIRE"),
                which(combined.diss.clim.common$SPP == "PIRI"),
                which(combined.diss.clim.common$SPP == "PITA")
                )
      
      Larix <- c(
                which(combined.diss.clim.common$SPP == "LALA")
                )
      
      Picea <- c(
                which(combined.diss.clim.common$SPP == "PIGL"),
                which(combined.diss.clim.common$SPP == "PIMA"),
                which(combined.diss.clim.common$SPP == "PIRU")
                )

      Abies <- c(
              which(combined.diss.clim.common$SPP == "ABBA")
              )
      
Juniperus <- c(
              which(combined.diss.clim.common$SPP == "JUVI")
              )

    Thuja <- c(
              which(combined.diss.clim.common$SPP == "THOC2")
              )
    
    Tsuga <- c(
              which(combined.diss.clim.common$SPP == "TSCA")
              )
    
    
    
    ### Set total dataframe for Genus
    
    combined.diss.clim.common$genus <- NA
    combined.diss.clim.common$genus[Pinus] <- "Pinus"
    combined.diss.clim.common$genus[Larix] <- "Larix"
    combined.diss.clim.common$genus[Picea] <- "Picea"
    combined.diss.clim.common$genus[Abies] <- "Abies"
    combined.diss.clim.common$genus[Thuja] <- "Thuja"
    combined.diss.clim.common$genus[Tsuga] <- "Tsuga"
    combined.diss.clim.common$genus[Juniperus] <- "Juniperus"
    combined.diss.clim.common$genus[Quercus] <- "Quercus"
    combined.diss.clim.common$genus[Fraxinus] <- "Fraxinus"
    combined.diss.clim.common$genus[Ailanthus] <- "Ailanthus"
    combined.diss.clim.common$genus[Robinia] <- "Robinia"
    combined.diss.clim.common$genus[Sassafras] <- "Sassafras"
    combined.diss.clim.common$genus[Acer] <- "Acer"
    combined.diss.clim.common$genus[Aesculus] <- "Aesculus"
    combined.diss.clim.common$genus[Betula] <- "Betula"
    combined.diss.clim.common$genus[Carpinus] <- "Carpinus"
    combined.diss.clim.common$genus[Cornus] <- "Cornus"
    combined.diss.clim.common$genus[Liquidambar] <- "Liquidambar"
    combined.diss.clim.common$genus[Liriodendron] <- "Liriodendron"
    combined.diss.clim.common$genus[Magnolia] <- "Magnolia"
    combined.diss.clim.common$genus[Platanus] <- "Platanus"
    combined.diss.clim.common$genus[Fagus] <- "Fagus"
    combined.diss.clim.common$genus[Nyssa] <- "Nyssa"
    combined.diss.clim.common$genus[Oxydendrum] <- "Oxydendrum"
    combined.diss.clim.common$genus[Tilia] <- "Tilia"
    combined.diss.clim.common$genus[Populus] <- "Populus"
    combined.diss.clim.common$genus[Prunus] <- "Prunus"
    combined.diss.clim.common$genus[Carya] <- "Carya"
    combined.diss.clim.common$genus[Cercis] <- "Cercis"
    combined.diss.clim.common$genus[Diospyros] <- "Diospyros"
    combined.diss.clim.common$genus[Juglans] <- "Juglans"
    combined.diss.clim.common$genus[Ulmus] <- "Ulmus"
    
  ###Add Family
  Aceraceae <- c(
              which(combined.diss.clim.common$genus == "Acer")
              )
  
  Betulaceae <- c(
                which(combined.diss.clim.common$genus == "Betula"),
                which(combined.diss.clim.common$genus == "Carpinus")
                )
  
  Cornaceae <- c(which(combined.diss.clim.common$genus == "Cornus"),
                 which(combined.diss.clim.common$genus == "Nyssa"))
  
  Cupressaceae <- c(which(combined.diss.clim.common$genus == "Thuja"),
                    which(combined.diss.clim.common$genus == "Juniperus"))
  
  Ebenaceae <- c(which(combined.diss.clim.common$genus == "Diospyros"))
  
  Ericaceae <- c(which(combined.diss.clim.common$genus == "Oxydendrum"))
  
  Fabaceae <- c(which(combined.diss.clim.common$genus == "Robinia"),
                    which(combined.diss.clim.common$genus == "Cercis"))
  
  Fagaceae <- c(which(combined.diss.clim.common$genus == "Quercus"),
                which(combined.diss.clim.common$genus == "Fagus"))
  
  Hamamelidaceae <- c(which(combined.diss.clim.common$genus == "Liquidambar"))
  
  Hippocastanaceae <- c(which(combined.diss.clim.common$genus == "Aesculus"))
  
  Juglandaceae <- c(which(combined.diss.clim.common$genus == "Carya"),
                    which(combined.diss.clim.common$genus == "Juglans"))
  
  Lauraceae <- c(which(combined.diss.clim.common$genus == "Sassafras"))
  
  Magnoliaceae <- c(which(combined.diss.clim.common$genus == "Liriodendron"),
                    which(combined.diss.clim.common$genus == "Magnolia"))
  
  Oleaceae <- c(which(combined.diss.clim.common$genus == "Fraxinus"))
  
  Pinaceae <- c(which(combined.diss.clim.common$genus == "Pinus"),
                which(combined.diss.clim.common$genus == "Larix"),
                which(combined.diss.clim.common$genus == "Picea"),
                which(combined.diss.clim.common$genus == "Abies"),
                which(combined.diss.clim.common$genus == "Tsuga")
                )
  
  Platanaceae <- c(which(combined.diss.clim.common$genus == "Platanus"))
  
  Rosaceae <- c(which(combined.diss.clim.common$genus == "Prunus"))
  
  Salicaceae <- c(which(combined.diss.clim.common$genus == "Populus"))
  
  Simaroubaceae <- c(which(combined.diss.clim.common$genus == "Ailanthus"))
  
  Tiliaceae <- c(which(combined.diss.clim.common$genus == "Tilia"))
  
  Ulmaceae <- c(which(combined.diss.clim.common$genus == "Ulmus"))
  
  #Make single genus-species column
    combined.diss.clim.common$GenusSpecies <- NA
    combined.diss.clim.common$GenusSpecies <- with(combined.diss.clim.common, paste(genus, species))
  
  #Set total dataframe for Family
  combined.diss.clim.common$family <- NA
    combined.diss.clim.common$family[Aceraceae] <- "Aceraceae"
    combined.diss.clim.common$family[Betulaceae] <- "Betulaceae"       
    combined.diss.clim.common$family[Cornaceae] <- "Cornaceae"
    combined.diss.clim.common$family[Cupressaceae] <- "Cupressaceae"
    combined.diss.clim.common$family[Ebenaceae] <- "Ebenaceae"
    combined.diss.clim.common$family[Ericaceae] <- "Ericaceae"
    combined.diss.clim.common$family[Fabaceae] <- "Fabaceae"
    combined.diss.clim.common$family[Fagaceae] <- "Fagaceae"
    combined.diss.clim.common$family[Hamamelidaceae] <- "Hamamelidaceae"
    combined.diss.clim.common$family[Hippocastanaceae] <- "Hippocastanaceae"
    combined.diss.clim.common$family[Juglandaceae] <- "Juglandaceae"
    combined.diss.clim.common$family[Sassafras] <- "Lauraceae"
    combined.diss.clim.common$family[Magnoliaceae] <- "Magnoliaceae"
    combined.diss.clim.common$family[Oleaceae] <- "Oleaceae"   
    combined.diss.clim.common$family[Pinaceae] <- "Pinaceae"
    combined.diss.clim.common$family[Platanaceae] <- "Platanaceae"
    combined.diss.clim.common$family[Rosaceae] <- "Rosaceae"
    combined.diss.clim.common$family[Salicaceae] <- "Salicaceae"
    combined.diss.clim.common$family[Simaroubaceae] <- "Simaroubaceae"
    combined.diss.clim.common$family[Tiliaceae] <- "Tiliaceae"
    combined.diss.clim.common$family[Ulmaceae] <- "Ulmaceae"

    #Order
    Sapindales <- c(which(combined.diss.clim.common$family == "Aceraceae"),
                    which(combined.diss.clim.common$family == "Hippocastanaceae"),
                    which(combined.diss.clim.common$family == "Simaroubaceae"))
    
    Fagales <- c(which(combined.diss.clim.common$family == "Betulaceae"),
                 which(combined.diss.clim.common$family == "Fagaceae"))
    
    Cornales <- c(which(combined.diss.clim.common$family == "Cornaceae"))
    
    Pinales <- c(which(combined.diss.clim.common$family == "Cupressaceae"),
                 which(combined.diss.clim.common$family == "Pinaceae"))
    
    Ebenales <- c(which(combined.diss.clim.common$family == "Ebenaceae"))
    Ericales <- c(which(combined.diss.clim.common$family == "Ericaceae"))
    Fabales <- c(which(combined.diss.clim.common$family == "Fabaceae"))
    
    Hamamelidales <- c(which(combined.diss.clim.common$family == "Hamamelidaceae"),
                       which(combined.diss.clim.common$family == "Platanaceae"))
    
    Juglandales <- c(which(combined.diss.clim.common$family == "Juglandaceae"))
    Laurales <- c(which(combined.diss.clim.common$family == "Lauraceae"))
    Magnoliales <- c(which(combined.diss.clim.common$family == "Magnoliaceae"))
    Scrophulariales <- c(which(combined.diss.clim.common$family == "Oleaceae"))
    Rosales <- c(which(combined.diss.clim.common$family == "Rosaceae"))
    Salicales <- c(which(combined.diss.clim.common$family == "Salicaceae"))
    Malvales <- c(which(combined.diss.clim.common$family == "Tiliaceae"))
    Urticales <- c(which(combined.diss.clim.common$family == "Ulmaceae"))
    
    #Set total dataframe for order
    combined.diss.clim.common$order <- NA
    combined.diss.clim.common$order[Sapindales] <- "Sapindales"
    combined.diss.clim.common$order[Fagales] <- "Fagales"       
    combined.diss.clim.common$order[Cornales] <- "Cornales"
    combined.diss.clim.common$order[Pinales] <- "Pinales"
    combined.diss.clim.common$order[Ebenales] <- "Ebenales"
    combined.diss.clim.common$order[Ericales] <- "Ericales"
    combined.diss.clim.common$order[Fabales] <- "Fabales"
    combined.diss.clim.common$order[Hamamelidales] <- "Hamamelidales"
    combined.diss.clim.common$order[Juglandales] <- "Juglandales"
    combined.diss.clim.common$order[Laurales] <- "Laurales"
    combined.diss.clim.common$order[Magnoliales] <- "Magnoliales"
    combined.diss.clim.common$order[Scrophulariales] <- "Scrophulariales"
    combined.diss.clim.common$order[Rosales] <- "Rosales"
    combined.diss.clim.common$order[Salicales] <- "Salicales"   
    combined.diss.clim.common$order[Malvales] <- "Malvales"
    combined.diss.clim.common$order[Urticales] <- "Urticales"

    #Subclass
    Rosidae <- c(which(combined.diss.clim.common$order == "Sapindales"),
                 which(combined.diss.clim.common$order == "Cornales"),
                 which(combined.diss.clim.common$order == "Fabales"),
                 which(combined.diss.clim.common$order == "Rosales"))
    
    Hamamelididae <- c(which(combined.diss.clim.common$order == "Fagales"),
                       which(combined.diss.clim.common$order == "Hamamelidales"),
                       which(combined.diss.clim.common$order == "Juglandales"),
                       which(combined.diss.clim.common$order == "Urticales"))
    
    Pinopsida <- c(which(combined.diss.clim.common$order == "Pinales"))
    
    Dilleniidae <- c(which(combined.diss.clim.common$order == "Ebenales"),
                     which(combined.diss.clim.common$order == "Ericales"),
                     which(combined.diss.clim.common$order == "Salicales"),
                     which(combined.diss.clim.common$order == "Malvales"))
    
    Magnoliidae <- c(which(combined.diss.clim.common$order == "Magnoliales"),
                     which(combined.diss.clim.common$order == "Laurales"))
    
    Asteridae <- c(which(combined.diss.clim.common$order == "Scrophulariales"))
    
        #Set total dataframe for Subclass
    combined.diss.clim.common$subclass <- NA
    combined.diss.clim.common$subclass[Rosidae] <- "Rosidae"
    combined.diss.clim.common$subclass[Hamamelididae] <- "Hamamelididae"       
    combined.diss.clim.common$subclass[Pinopsida] <- "Pinopsida"
    combined.diss.clim.common$subclass[Dilleniidae] <- "Dilleniidae"
    combined.diss.clim.common$subclass[Magnoliidae] <- "Magnoliidae"
    combined.diss.clim.common$subclass[Asteridae] <- "Asteridae"


combined.diss.clim.common$GenusSpecies <- factor(combined.diss.clim.common$GenusSpecies, 
                  levels = c("Juniperus virginiana","Thuja occidentalis","Abies balsamea","Tsuga canadensis","Picea mariana", "Picea rubens","Pinus strobus","Pinus palustris","Pinus taeda","Pinus rigida","Sassafras albidum","Magnolia virginiana","Liriodendron tulipifera","Liquidambar styraciflua","Oxydendrum arboreum","Nyssa sylvatica","Fraxinus americana","Fraxinus nigra","Tilia americana","Acer rubrum","Acer saccharum","Acer negundo","Populus grandidentata","Populus tremuloides","Prunus serotina","Ulmus americana","Betula lenta","Betula papyrifera","Betula alleghaniensis","Juglans nigra","Carya glabra","Carya cordiformis","Carya ovata","Carya tomentosa","Fagus grandifolia","Quercus laurifolia","Quercus alba","Quercus velutina","Quercus nigra","Quercus laevis","Quercus montana","Quercus rubra","Quercus virginiana","Quercus coccinea","Quercus muehlenbergii","Quercus phellos"), ordered = TRUE) #Quercus coccinea added after tree made need to figure out where it actually belongs

combined.diss.clim.common$WOOD <- factor(combined.diss.clim.common$WOOD, levels = c("large.resin","small.resin", "no.resin","ring.porous","semiring.porous","diffuse.porous"), ordered = TRUE)

combined.diss.clim.common$PFT <- factor(combined.diss.clim.common$PFT, levels = c("Early.Conifer","Late.Conifer","Early.Hardwood","Mid.Hardwood","Late.Hardwood"), ordered = TRUE)

write.csv(x = combined.diss.clim.common, file = "/Users/josh/Dropbox/Dissertation/CH1_Treerings/Data/combined.diss.clim.common.wood.phyl.csv")

#manually moved columns
combined.diss.clim.common.order <- read.csv("/Users/josh/Dropbox/Dissertation/CH1_Treerings/Data/combined.diss.clim.common.wood.phyl.csv")

colnames(combined.diss.clim.common.order)[104:153] <- colnames.yrs#colnames(combined.dmg.clean) #get rid of stupid "x" before year

#Reformat data so I can do linear models for more detailed analysis
# #recreate data frame to be compliant with structure needed to include time as a variable (50 years)
ids.rep.new.diss.common.order <- data.frame(matrix(NA,nrow=1,ncol=103))
colnames(ids.rep.new.diss.common.order) <- colnames(combined.diss.clim.common.order[1:103])
for(i in 1:nrow(combined.diss.clim.common.order)){
  ids.rep.diss.common.order <- combined.diss.clim.common.order[rep(i,times=50),1:103]
  ids.rep.new.diss.common.order <- rbind(ids.rep.new.diss.common.order,ids.rep.diss.common.order)
}
#remove row of NAs
combined.ids.diss.common.order <- ids.rep.new.diss.common.order[2:nrow(ids.rep.new.diss.common.order),]

#transpose and concatenate increments for all trees
inc.new.diss.common.order <- vector("numeric",length=0)
for(i in 1:nrow(combined.diss.clim.common.order)){
  inc.diss.common.order <- as.matrix(combined.diss.clim.common.order[i,104:153])
  inc.new.diss.common.order <- as.numeric(c(inc.new.diss.common.order,inc.diss.common.order))
}

#add ring increments
combined.ids.incs.diss <- cbind(combined.ids.diss.common.order,inc.new.diss.common.order)

#rename inc variable
setnames(combined.ids.incs.diss,old='inc.new.diss.common.order',new='INCREMENT')

YEAR <- seq(1967,2016,1)
YEAR <- rep(YEAR,1568)

#final version with 50yrs of growth
combined.all.diss.common <- cbind(combined.ids.incs.diss,YEAR)

#ANALYSIS
lm.space <- lmer(INCREMENT ~ YEAR + (1|SITE/PLOT/SUB/TAG), data = combined.all.diss.common)
summary(lm.space)

lm.clim.trait.spp <- lmer(INCREMENT ~ YEAR + (1|SPP) + (1|TEMP_K) +(1|PRECIP) + (1|PFT) + (1|WOOD) + (1|SITE/PLOT/SUB/TAG), data = combined.all.diss.common)
summary(lm.clim.trait.spp)


summary(nsc.kitchen.lm)
plot((nsc.kitchen.lm))
```
You can also embed plots, for example:

```{r fig.width=7, fig.height=6}
plot(cars)
```


