Tree ring bayesian hierarchical mixed effects model
========================================================

The tree ring analysis to-date has included a hierarchical structure to include individuals, subplots, plots and sites and some individual effects like species. I am now adding year x plot and year x site interactions to see how much variance is explained by climate at each site and within a site.

You can embed an R code chunk like this:

```{r}
#Required Packages
library(rjags)
library(data.table)

#Source required PEcAn scripts for reading and matching treerings and plot data
source("/Users/Josh/Documents/Research/pecan/modules/data.land/R/Read_Tuscon.R")
source("/Users/Josh/Documents/Research/pecan/modules/data.land/R/matchInventoryRings.R")
```

Using most up-to-date version of adult census data for treering matching. Includes 2015 census data, but not boolean damage column. All other columns should be the same as analysis done for AGU and EE seminar.

```{r}
#read Tucson format ringwidth files
trees.diss <- read.csv("/Users/josh/Dropbox/Dissertation/CH1_Treerings/Data/Adult_Field_Data_JAM_MCD.csv")
rings.diss <- Read_Tuscon("/Users/josh/Dropbox/Dissertation/CH1_Treerings/Data/Windendro_text/All_Sites")

#Match Tucson ringwidth files and adult census data
combined.diss <- matchInventoryRings(trees=trees.diss,rings.diss)

```

```{r}
#Example creating interaction term
Y = seq(1,10)
L = as.vector(x=c("A","B","C"))
YL = expand.grid(Y,L) #creates matrix of all combinations as characters
YL.vector = as.vector(paste0(YL$Var1,YL$Var2)) #vector of all combinations as characters
#Can now use this to create interaction term for years x plot and years x site

```

You can also embed plots, for example:

```{r fig.width=7, fig.height=6}
plot(cars)
```


